NMRlipids IV: Headgroup \& glycerol backbone structures, and cation binding in bilayers with PE and PG lipids

### 2019-04-17 Manuscript/manuscriptPGPE.tex
Add PG and PE data committed by Antonio Peon https://github.com/NMRLipids/MATCH/pull/59 and others, update figures and manuscript

\begin{table*}[htb]
  %\begin{sidewaystable*}[!p]
  \centering
  \caption{List of single lipid type MD simulations without additional ions.
    % The salt concentrations calculated as [salt]=N$_{\rm c} \times$[water]\,/\,N$_{\rm w}$, where [water]\,=\,55.5~M.
    % these correspond the concentrations reported in the experiments by Akutsu et al.~\cite{akutsu81}.
    % The lipid force fields named as in our previous work~\cite{botan15}.
  }\label{systems}
  \begin{minipage}[t]{\textwidth}
    \begin{tabular}{l c c r r r r r r c c}
      %\hline
      % some footnotes are not visible in typeset-MS (pdf)
      lipid/counter-ions & force field for lipids / ions & \footnote{Number of lipid molecules with largest mole fraction}N$_{\rm l}$   &  \footnote{Number of water molecules}N$_{\rm w}$    & \footnote{Simulation temperature}T (K)  & \footnote{Total simulation time}t$_{{\rm sim}}$(ns) & \footnote{Time used for analysis}t$_{{\rm anal}}$ (ns) &   \footnote{Reference for simulation files}files\\
      \hline
      DPPE  & Slipids \cite{jambeck12b}    & 288 	& 9386  & 336  & 200 & 100 & \cite{slipidsDPPEfiles}  \\
      DPPE  & GROMOS-CKP    \cite{??}      & 128	& ?     & 342  & 2$\times$500 & 2$\times$400 & \cite{gromosCKPdppe} \todoi{Not analyzed yet, waiting for the code for UA simulations.}  \\
      \hline
      POPE  & GROMOS 43A1-S3 \cite{??}     & 128	& ?     & 313  & 2$\times$200 & 2$\times$100 & \cite{gromos43a1s3POPEfiles} \todoi{Not analyzed yet, waiting for the code for UA simulations.}  \\
      POPE  & CHARMM36ua \cite{??}         & 336	& ?     & 310  & 2$\times$200 & 2$\times$100 & \cite{charmm36uaPOPEfiles} \todoi{Not analyzed yet, waiting for the code for UA simulations.}  \\
      POPE  & Slipids \cite{jambeck12b,??} & 336	& ?     & 310  & 2$\times$200 &  2$\times$100 & \cite{slipidsPOPEfiles}  \\
      POPE  & OPLS-UA vdW on H \cite{??}   & 128	& ?     & 303  & 2$\times$200 & 2$\times$100 & \cite{OPLSuaWvdWPOPEfiles} \todoi{Not analyzed yet, waiting for the code for UA simulations.}  \\
      POPE  & OPLS-UA \cite{??}            & 128	& ?     & 303  & 2$\times$200 & 2$\times$100 & \cite{OPLSuaPOPEfiles} \todoi{Not analyzed yet, waiting for the code for UA simulations.}  \\
      POPE  & Berger-based \cite{??}       & 128	& ?     & 303  & 2$\times$200 & 2$\times$100 & \cite{bergerPOPEfiles} \todoi{Not analyzed yet, waiting for the code for UA simulations.}  \\
      POPE  & Berger-based2 \cite{??}      & 128	& ?     & 303  & 2$\times$200 & 2$\times$100 & \cite{berger2POPEfiles} \todoi{Not analyzed yet, waiting for the code for UA simulations.}  \\
      POPE  & GROMOS-CKP    \cite{??}      & 128	& ?     & 313  & 2$\times$500 & 2$\times$400 & \cite{gromosCKPpope} \todoi{Not analyzed yet, waiting for the code for UA simulations.}  \\
      \hline
      DOPE  & Berger-based \cite{??}       & 128	& ?     & 271  & 2$\times$200 & 2$\times$100 & \cite{bergerDOPEfiles} \todoi{Not analyzed yet, waiting for the code for UA simulations.}  \\
      DOPE  & Berger-based2 \cite{??}      & 128	& ?     & 271  & 2$\times$300 & 2$\times$100 & \cite{berger2DOPEfiles} \todoi{Not analyzed yet, waiting for the code for UA simulations.}  \\
      DOPE  & GROMOS-CKP    \cite{??}      & 128	& ?     & 271  & 2$\times$500 & 2$\times$400 & \cite{gromosCKPdope} \todoi{Not analyzed yet, waiting for the code for UA simulations.}  \\
      \hline
      POPG/K$^+$  & CHARMM36 \cite{??} \todoi{Correct citation for CHARMM POPG}   & 118& 4110  & 298  & 100 & 100 & \cite{CHARMM36popg}  \\
      POPG/Na$^+$  & Slipids \cite{jambeck13}   & 288 	& 10664  & 298  & 250 & 100 & \cite{slipidsPOPGfiles} \\
      \hline
      DPPG/Na$^+$  & Slipids \cite{jambeck13}   & 288 	& 11232  & 314  & 200 & 100 & \cite{slipidsDPPGfiles} \\
      DPPG/Na$^+$  & Slipids \cite{jambeck13}   & 288 	& 11232  & 298  & 400 & 100 & \cite{slipidsDPPGfilesT298K} \\
    \end{tabular}
  \end{minipage}
  %\end{sidewaystable*} 
\end{table*}

\begin{table*}[htb]
  %\begin{sidewaystable*}[!p]
  \centering
  \caption{List of MD simulations with lipid mixtures and additional ions.
    % The salt concentrations calculated as [salt]=N$_{\rm c} \times$[water]\,/\,N$_{\rm w}$, where [water]\,=\,55.5~M.
    % these correspond the concentrations reported in the experiments by Akutsu et al.~\cite{akutsu81}.
    % The lipid force fields named as in our previous work~\cite{botan15}.
  }\label{systems}
  \begin{minipage}[t]{\textwidth}
    \begin{tabular}{l c c r r r r r r c c}
      %\hline
      % some footnotes are not visible in typeset-MS (pdf)
      lipid/counter-ions & force field for lipids / ions & NaCl (mM) & CaCl$_2$\,(M) &  \footnote{Number of lipid molecules with largest mole fraction}N$_{\rm l}$   &  \footnote{Number of water molecules}N$_{\rm w}$   & \footnote{Number of additional cations}N$_{\rm c}$  & \footnote{Simulation temperature}T (K)  & \footnote{Total simulation time}t$_{{\rm sim}}$(ns) & \footnote{Time used for analysis}t$_{{\rm anal}}$ (ns) &   \footnote{Reference for simulation files}files\\
      \hline
      POPC:POPG (1:1)/K$^+$  & CHARMM36 \cite{??}        &0          & 0      & ? & ? & ?  &  ?  & ? & ? & \cite{??} \todoi{Data to be uploaded by J. Madsen. Details to be filled once we have the data}  \\
      POPC:POPG (1:1)/K$^+$  & CHARMM36 \cite{??}        &0          & 0.15 \todoi{Concentration to be checked.}  & ? & ? & ?  &  ?  & ? & ? & \cite{??} \todoi{Data to be uploaded by J. Madsen. Details to be filled once we have the data}  \\
      POPC:POPG (1:1)/K$^+$  & CHARMM36 \cite{??}        &0          & 1.0  \todoi{Concentration to be checked.}  & ? & ? & ?  &  ?  & ? & ? & \cite{??} \todoi{Data to be uploaded by J. Madsen. Details to be filled once we have the data}  \\
      POPC:POPG (4:1)/K$^+$  & CHARMM36 \cite{??}        &0          & 0      & ? & ? & ?  &  ?  & ? & ? & \cite{??} \todoi{Data to be uploaded by J. Madsen. Details to be filled once we have the data}  \\
      POPC:POPG (4:1)/K$^+$  & CHARMM36 \cite{??}        &0          & 0.15 \todoi{Concentration to be checked.}  & ? & ? & ?  &  ?  & ? & ? & \cite{??} \todoi{Data to be uploaded by J. Madsen. Details to be filled once we have the data}  \\
      POPC:POPG (4:1)/K$^+$  & CHARMM36 \cite{??}        &0          & 1.0  \todoi{Concentration to be checked.}  & ? & ? & ?  &  ?  & ? & ? & \cite{??} \todoi{Data to be uploaded by J. Madsen. Details to be filled once we have the data}  \\
      \hline
      POPC             & CHARMM36 \cite{??}        &0          & 0  & 256 & 8704 & 0  &  300  & 300 & 250 & \cite{POPCcharmm300K} \\
      POPC:POPE (1:1)  & CHARMM36 \cite{??}        &0          & 0  & 128 & 8704 & 0  &  300  & 300 & 250 & \cite{POPC1POPE1charmm36} \\
     \hline
      POPC             & Berger \cite{??} \todoi{This is probable not plain berger, correct force filed should be described.}  &0  & 0  & 256 & 10240 & 0  &  300  & 300 & 200 & \cite{POPCberger300K} \\
      POPC:POPE (1:1)  & Berger \cite{??}  \todoi{This is probable not plain berger, correct force filed should be described.} &0          & 0  & 128 & 11008 & 0  &  300  & 300 & 200 & \cite{POPC1POPE1berger} \\
      POPC:DOPE (1:1)  & Berger \cite{??}  \todoi{This is probable not plain berger, correct force filed should be described.}         &0          & 0  & 128 & 10240 & 0  &  300  & 300 & 200 & \cite{POPC1DOPE1berger} \\
      DOPC             & Berger \cite{??}  \todoi{This is probable not plain berger, correct force filed should be described.}         &0          & 0  & 256 & 11008 & 0  &  300  & 300 & 200 & \cite{DOPCberger300K} \\
      DOPC:DOPE (1:1)  & Berger \cite{??}   \todoi{This is probable not plain berger, correct force filed should be described.}        &0          & 0  & 128 & 11008 & 0  &  300  & 300 & 200 & \cite{DOPC1DOPE1berger} \\
    \end{tabular}
  \end{minipage}
  %\end{sidewaystable*} 
\end{table*}

@@

\subsection{PE headgroup and glycerol backbone}
\todo{Add results from this discussion: \url{http://nmrlipids.blogspot.com/2017/03/nmrlipids-iv-headgroup-glycerol.html?showComment=1502290977005#c4964787551006026647}}
Order parameters from Slipids simulations 
and experiments for DPPE are shown in Fig. \ref{HGorderParametersPE}.
Glycerol backbone order parameters in Slipids are off from experiments,
as already observed previously for PC lipids \cite{botan15}.
Order parameter signs for PE are not experimentally measured yet.
For headgroup the signs are set to give best agreement with simulations
and for glycerol to be consistent with experimental signs for PC.
Order parameter for $\beta$ carbon shows apparent agreement
with experiments. However, the sign of beta order parameter is positive,
in contrast to PC where negative sign was measured. Thus, the the beta order
parameter agrees with experiment with the assumption that its sign is opposite
than for PC. This is yet to be confirmed by experiments.
Order parameter for $\alpha$ carbon is too close to zero, even if the sign
would be correct.

\begin{figure}[]
  \centering
  \includegraphics[width=9.0cm]{../Figs/HGorderparametersPE.pdf}
  \caption{\label{HGorderParametersPE}
    Order parameters for DPPE headgroup and glycerol
    backbone from simulations with Slipids \cite{??} and experiments
    (DPPE from \cite{seelig76} and E.coliPE from \cite{gally81}).
    Absolute values are shown, because signs are not known experimentally.
  }
  \todo{Experimental signs of the order parameters would be highly useful.} \\
\end{figure}

@@

\subsection{PG headgroup and glycerol backbone}
Comparison between experiments and simulations for PG lipids is shown
in Fig. \ref{HGorderParametersPOPG}. The signs are not yet measured
experimentally. They are set to give the best argeement with experiments.
This would suggest that the $\beta$ order parameter would be positive,
in contrast to PC and PS headgroups, were negative signs were measured.
Even thought the signs turned out to be correct, the tested models would
not give a very good argeement with the experiments.
\begin{figure}[]
  \centering
  \includegraphics[width=9.0cm]{../Figs/HGorderparametersPG.pdf}
  \caption{\label{HGorderParametersPOPG}
    Order parameters for PG headgroup and glycerol backbone from simulations and experiments without CaCl$_2$ 
    (POPG from \cite{borle85} contains 10mM of PIPES, DPPG from \cite{wohlgemuth80} contains 10mM PIPES and 100mM CaCl,
    E.Coli PG results from \cite{gally81}).
    Signs are not known for experimental data. They are determined to give best agreement
    with simulations. This is not reliable and should be corrected when experimental
    data becomes available.
  }
  \todo{More simulation data for lipids with different headgroups to be collected} \\
  \todo{CHARMM GUI simulation contains only counter ions as potassium.
    All experiments here contain some amount of sodium salt. The best
    ion concentrations for comparison should be figured out.} \\
  \todo{Experimental signs of the order parameters would highly useful.}
\end{figure}

@@

\subsection{Lipid headgroup interactions in PC:PE and PC:PG mixtures}
The headgroup 
order parameters increase with the addition of negatively charged PS and PG lipids,
decrease when mixed with positively charged surfactants and
are less affected by the addition of zwitterionic PE lipids or cholesterol.
In addition to the results summarized in Fig. \ref{HGorderparametersPCvsPEPSPGchol},
also mixtures of PC with negatively charged PI, CL, PA, and zwitterionic SM
follow the eletrometer concept \cite{scherer87}.
to mixtures of differently charged
lipids are collected from different experiments in Fig. \ref{HGorderparametersPCvsPEPSPGchol}.
%\section{PC lipid headgroup response to different mixtures in experiments}
As shown in Fig. \ref{HGorderparametersPCvsPEPSPGchol}, order parameters of PC
headgroup behave in various lipid mixtures as expected from the electrometer concept \cite{seelig87, scherer87},
i.e., order parameters increase when anionic lipids are mixed with PC and decrease with cationic
surfactants. The changes with the addition of neutral lipids is significantly smaller.
\begin{figure}[]
  \centering
  \includegraphics[width=8.0cm]{../Figs/HGorderparametersPCvsPEPSPGchol.eps}
  \caption{\label{HGorderparametersPCvsPEPSPGchol}
    PC headgroup order parameters from experiments of mixtures with
    PE, PS, PG and cholesterol \cite{scherer87,scherer89,ferreira13}.
    Signs are determined as discussed in \cite{botan15,ollila16}.
  }
\end{figure}

@@

\begin{figure*}[]
  \centering
  \includegraphics[width=16.0cm]{../Figs/HGorderparametersPCvsPEPSPG.eps}
  \caption{\label{HGorderparametersPCvsPEPSPG}
    PC headgroup order parameters from mixtures with PE, PS and PG
    lipids with various mole fractions from different simulation models and experiments \cite{scherer87}.
    Signs are determined as discussed in \cite{botan15,ollila16}.
  }
  \todo{Simulation of CHARMM36 at 298K should be maybe rerun with Gromacs 5.}
\end{figure*}

@@

@@

Comparison of Ca$2+$ binding in PG between CHARMM36 simulations and experiments \cite{borle85}
is shown in Fig. \ref{changesWITHCaClPG}. The decrease of $\alpha$ order parameter
is in agreement with experiments, while decerase of $\beta$ order parameter is overestimated.
The result is very similar to the results with PC in NMRlipids II publication \cite{catte16}.
It should be, however, noted that the $\beta$-order parameters are not actually measured for PG,
but they are calculated from empirical relation $\Delta S_{\beta}=0.43\Delta S_{\alpha}$ \cite{akutsu81}.
Anyway, the data presented in NMRlipids II project and in Fig. \ref{changesWITHCaClPG} together
suggest that Calcium binding is similarly overestimated by CHARMM36 model in pure POPC bilayers and mixtures with
POPG. The good agreement of $\alpha$ carbon would be explained by too weak dependence of its order
parameter of bound charge
\todo{The response of CHARMM36 to cationic surfactant against experiments \cite{scherer89} to be checked.
  I have already ran the simulations, analysis to be done.}.
\begin{figure}[]
  \centering
  \includegraphics[width=9.0cm]{../Figs/CHANGESwithCaClPG.eps}
  \caption{\label{changesWITHCaClPG}
    PG order parameters as a function CaCl$_2$ concentration from experiments \cite{borle85} and CHARMM36 simulations.
    Note that beta order parameter is calculated from empirical relation $\Delta S_{\beta}=0.43\Delta S_{\alpha}$ \cite{akutsu81}, not actually measured. 
  }
\end{figure}

@@

Also the experimental order parameters for PS and PG headgroups
as a function of CaCl$_2$ concentration are shown in Fig. \ref{PSPGchangesWITHCaCl}.
\todo{These should be compared to simulations for potential structural interpretation of the changes.}
\begin{figure}[]
  \centering
  \includegraphics[width=9.0cm]{../Figs/PGwithCaCl.eps}
  \caption{\label{PSPGchangesWITHCaCl}
    PG and PS order parameters a function CaCl$_2$ concentration taken from \cite{borle85} and \cite{roux90}, respectively.
  }
  \todo{Get the small concentration data from the inserts}
\end{figure}

>>>>

\begin{table*}[htb]
  %\begin{sidewaystable*}[!p]
  \centering
  \caption{List of single lipid type MD simulations without additional ions.
    % The salt concentrations calculated as [salt]=N$_{\rm c} \times$[water]\,/\,N$_{\rm w}$, where [water]\,=\,55.5~M.
    % these correspond the concentrations reported in the experiments by Akutsu et al.~\cite{akutsu81}.
    % The lipid force fields named as in our previous work~\cite{botan15}.
  }\label{systems}
  \begin{minipage}[t]{\textwidth}
    \begin{tabular}{l c c r r r r r r c c}
      %\hline
      % some footnotes are not visible in typeset-MS (pdf)
      lipid/counter-ions & force field for lipids / ions & \footnote{Number of lipid molecules with largest mole fraction}N$_{\rm l}$   &  \footnote{Number of water molecules}N$_{\rm w}$    & \footnote{Simulation temperature}T (K)  & \footnote{Total simulation time}t$_{{\rm sim}}$(ns) & \footnote{Time used for analysis}t$_{{\rm anal}}$ (ns) &   \footnote{Reference for simulation files}files\\
      \hline
      DPPE  & Slipids \cite{jambeck12b}    & 288 	& 9386  & 336  & 200 & 100 & \cite{slipidsDPPEfiles}  \\
      DPPE  & GROMOS-CKP    \cite{??}      & 128	& ?     & 342  & 2$\times$500 & 2$\times$400 & \cite{gromosCKPdppe} \todoi{Not analyzed yet, waiting for the code for UA simulations.}  \\
      \hline
      POPE  & GROMOS 43A1-S3 \cite{??}     & 128	& ?     & 313  & 2$\times$200 & 2$\times$100 & \cite{gromos43a1s3POPEfiles} \todoi{Not analyzed yet, waiting for the code for UA simulations.}  \\
      POPE  & CHARMM36ua \cite{??}         & 336	& ?     & 310  & 2$\times$200 & 2$\times$100 & \cite{charmm36uaPOPEfiles} \todoi{Not analyzed yet, waiting for the code for UA simulations.}  \\
      POPE  & CHARMM36 \cite{??}           & 144	& 5760  & 310  & 500          & 400          & \cite{charmm36POPEfiles} \\
      POPE  & Slipids \cite{jambeck12b,??} & 336	& ?     & 310  & 2$\times$200 &  2$\times$100 & \cite{slipidsPOPEfiles}  \\
      POPE  & OPLS-UA vdW on H \cite{??}   & 128	& ?     & 303  & 2$\times$200 & 2$\times$100 & \cite{OPLSuaWvdWPOPEfiles} \todoi{Not analyzed yet, waiting for the code for UA simulations.}  \\
      POPE  & OPLS-UA \cite{??}            & 128	& ?     & 303  & 2$\times$200 & 2$\times$100 & \cite{OPLSuaPOPEfiles} \todoi{Not analyzed yet, waiting for the code for UA simulations.}  \\
      POPE  & Berger-based \cite{??}       & 128	& ?     & 303  & 2$\times$200 & 2$\times$100 & \cite{bergerPOPEfiles} \todoi{Not analyzed yet, waiting for the code for UA simulations.}  \\
      POPE  & Berger-based2 \cite{??}      & 128	& ?     & 303  & 2$\times$200 & 2$\times$100 & \cite{berger2POPEfiles} \todoi{Not analyzed yet, waiting for the code for UA simulations.}  \\
      POPE  & GROMOS-CKP    \cite{??}      & 128	& ?     & 313  & 2$\times$500 & 2$\times$400 & \cite{gromosCKPpope} \todoi{Not analyzed yet, waiting for the code for UA simulations.}  \\
      \hline
      DOPE  & Berger-based \cite{??}       & 128	& ?     & 271  & 2$\times$200 & 2$\times$100 & \cite{bergerDOPEfiles} \todoi{Not analyzed yet, waiting for the code for UA simulations.}  \\
      DOPE  & Berger-based2 \cite{??}      & 128	& ?     & 271  & 2$\times$300 & 2$\times$100 & \cite{berger2DOPEfiles} \todoi{Not analyzed yet, waiting for the code for UA simulations.}  \\
      DOPE  & GROMOS-CKP    \cite{??}      & 128	& ?     & 271  & 2$\times$500 & 2$\times$400 & \cite{gromosCKPdope} \todoi{Not analyzed yet, waiting for the code for UA simulations.}  \\
      \hline
      POPG/K$^+$  & CHARMM36 \cite{??} \todoi{Correct citation for CHARMM POPG}   & 118& 4110  & 298  & 100 & 100 & \cite{CHARMM36popg}  \\
      POPG/Na$^+$  & Slipids \cite{jambeck13}   & 288 	& 10664  & 298  & 250 & 100 & \cite{slipidsPOPGfiles} \\
      \hline
      DPPG/Na$^+$  & Slipids \cite{jambeck13}   & 288 	& 11232  & 314  & 200 & 100 & \cite{slipidsDPPGfiles} \\
      DPPG/Na$^+$  & Slipids \cite{jambeck13}   & 288 	& 11232  & 298  & 400 & 100 & \cite{slipidsDPPGfilesT298K} \\
    \end{tabular}
  \end{minipage}
  %\end{sidewaystable*} 
\end{table*}

\begin{table*}[htb]
  %\begin{sidewaystable*}[!p]
  \centering
  \caption{List of MD simulations with lipid mixtures and additional ions.
    % The salt concentrations calculated as [salt]=N$_{\rm c} \times$[water]\,/\,N$_{\rm w}$, where [water]\,=\,55.5~M.
    % these correspond the concentrations reported in the experiments by Akutsu et al.~\cite{akutsu81}.
    % The lipid force fields named as in our previous work~\cite{botan15}.
  }\label{systems}
  \begin{minipage}[t]{\textwidth}
    \begin{tabular}{l c c r r r r r r c c}
      %\hline
      % some footnotes are not visible in typeset-MS (pdf)
      lipid/counter-ions & force field for lipids / ions & NaCl (M) & CaCl$_2$\,(M) &  \footnote{Number of lipid molecules with largest mole fraction}N$_{\rm l}$   &  \footnote{Number of water molecules}N$_{\rm w}$   & \footnote{Number of additional cations}N$_{\rm c}$  & \footnote{Simulation temperature}T (K)  & \footnote{Total simulation time}t$_{{\rm sim}}$(ns) & \footnote{Time used for analysis}t$_{{\rm anal}}$ (ns) &   \footnote{Reference for simulation files}files\\
      \hline
      POPC                   & CHARMM36 \cite{??}        & 0.11      & 0  & 500     & 25000 & 48  &  310  & 500 & 100 & \cite{POPCcharmm150mMNaCl301K}  \\
      POPC:POPG (7:3)        & CHARMM36 \cite{??}        & 0.11      & 0  & 350     & ?     & ?   &  310  & 500 & 100 & \cite{POPC7POPG1charmm36NaCl}  \\
      POPC:POPG (1:1)/K$^+$  & CHARMM36 \cite{??}        &0          & 0      & ? & ? & ?  &  ?  & ? & ? & \cite{??} \todoi{Data to be uploaded by J. Madsen. Details to be filled once we have the data}  \\
      POPC:POPG (1:1)/K$^+$  & CHARMM36 \cite{??}        &0          & 0.15 \todoi{Concentration to be checked.}  & ? & ? & ?  &  ?  & ? & ? & \cite{??} \todoi{Data to be uploaded by J. Madsen. Details to be filled once we have the data}  \\
      POPC:POPG (1:1)/K$^+$  & CHARMM36 \cite{??}        &0          & 1.0  \todoi{Concentration to be checked.}  & ? & ? & ?  &  ?  & ? & ? & \cite{??} \todoi{Data to be uploaded by J. Madsen. Details to be filled once we have the data}  \\
      POPC:POPG (4:1)/K$^+$  & CHARMM36 \cite{??}        &0          & 0      & ? & ? & ?  &  ?  & ? & ? & \cite{??} \todoi{Data to be uploaded by J. Madsen. Details to be filled once we have the data}  \\
      POPC:POPG (4:1)/K$^+$  & CHARMM36 \cite{??}        &0          & 0.15 \todoi{Concentration to be checked.}  & ? & ? & ?  &  ?  & ? & ? & \cite{??} \todoi{Data to be uploaded by J. Madsen. Details to be filled once we have the data}  \\
      POPC:POPG (4:1)/K$^+$  & CHARMM36 \cite{??}        &0          & 1.0  \todoi{Concentration to be checked.}  & ? & ? & ?  &  ?  & ? & ? & \cite{??} \todoi{Data to be uploaded by J. Madsen. Details to be filled once we have the data}  \\
      \hline
      POPC             & CHARMM36 \cite{??}        &0          & 0  & 256 & 8704 & 0  &  300  & 300 & 250 & \cite{POPCcharmm300K} \\
      POPC:POPE (1:1)  & CHARMM36 \cite{??}        &0          & 0  & 128 & 8704 & 0  &  300  & 300 & 250 & \cite{POPC1POPE1charmm36} \\
     \hline
      POPC                   & Slipid \cite{??}        & 0.11      & 0  & 500     & 25000 & 48  &  310  & 500 & 100 & \cite{POPCslipid150mMNaCl301K}  \\
      POPC:POPG (7:3)        & Slipid \cite{??}        & ?         & 0  & ?     & ?     & ?   &  310  & 500 & 100 & \cite{??} \todoi{Zenodo entry unclear.}  \\
     \hline
      POPC             & Berger \cite{??} \todoi{This is probable not plain berger, correct force filed should be described.}  &0  & 0  & 256 & 10240 & 0  &  300  & 300 & 200 & \cite{POPCberger300K} \\
      POPC:POPE (1:1)  & Berger \cite{??}  \todoi{This is probable not plain berger, correct force filed should be described.} &0          & 0  & 128 & 11008 & 0  &  300  & 300 & 200 & \cite{POPC1POPE1berger} \\
      POPC:DOPE (1:1)  & Berger \cite{??}  \todoi{This is probable not plain berger, correct force filed should be described.}         &0          & 0  & 128 & 10240 & 0  &  300  & 300 & 200 & \cite{POPC1DOPE1berger} \\
     \hline
      DOPC             & Berger \cite{??}  \todoi{This is probable not plain berger, correct force filed should be described.}         &0          & 0  & 256 & 11008 & 0  &  300  & 300 & 200 & \cite{DOPCberger300K} \\
      DOPC:DOPE (1:1)  & Berger \cite{??}   \todoi{This is probable not plain berger, correct force filed should be described.}        &0          & 0  & 128 & 11008 & 0  &  300  & 300 & 200 & \cite{DOPC1DOPE1berger} \\
      \hline
      POPG             & Slipids \cite{??} \todoi{Ion parameters?}       & 0.11          & 0  & 500 & 25000 & 49  &  310  & 500 & 100 & \cite{POPGslipids150mMNaCl} \\
      POPG             & CHARMM36 \cite{??}        & 0.11          & 0  & 500 & 25000 & 49  &  310  & 500 & 100 & \cite{POPGcharmm150mMNaCl} \\
     \hline
      POPE             & Slipids \cite{??} \todoi{Ion parameters?}        & 0.11          & 0  & 500 & 25000 & 50  &  310  & 500 & 100 & \cite{POPEslipids150mMNaCl} \\
      POPE             & CHARMM36 \cite{??}        & 0.11          & 0  & 500 & 25000 & 50  &  310  & 500 & 100 & \cite{POPEcharmm150mMNaCl} \\
    \end{tabular}
  \end{minipage}
  %\end{sidewaystable*} 
\end{table*}

@@

\subsection{PE headgroup and glycerol backbone}
%\todo{Add results from this discussion: \url{http://nmrlipids.blogspot.com/2017/03/nmrlipids-iv-headgroup-glycerol.html?showComment=1502290977005#c4964787551006026647}}
Order parameters from Slipids simulations 
and experiments for DPPE are shown in Fig. \ref{HGorderParametersPE}.
Glycerol backbone order parameters in Slipids are off from experiments,
as already observed previously for PC lipids \cite{botan15}.
Order parameter signs for PE are not experimentally measured yet.
For headgroup the signs are set to give best agreement with simulations
and for glycerol to be consistent with experimental signs for PC.
Order parameter for $\beta$ carbon shows apparent agreement
with experiments. However, the sign of beta order parameter is positive,
in contrast to PC where negative sign was measured. Thus, the the beta order
parameter agrees with experiment with the assumption that its sign is opposite
than for PC. This is yet to be confirmed by experiments.
Order parameter for $\alpha$ carbon is too close to zero, even if the sign
would be correct.

\begin{figure}[]
  \centering
  \includegraphics[width=9.0cm]{../Figs/HGorderparametersPE.pdf}
  \caption{\label{HGorderParametersPE}
    Order parameters for DPPE headgroup and glycerol
    backbone from simulations with Slipids \cite{??} and experiments
    (DPPE from \cite{seelig76} and E.coliPE from \cite{gally81}).
    Absolute values are shown, because signs are not known experimentally.
  }
  \todo{Results from united atom simulations yet to be added} \\
\end{figure}

@@

\subsection{PG headgroup and glycerol backbone}
Comparison between experiments and simulations for PG lipids is shown
in Fig. \ref{HGorderParametersPOPG}. The signs are not yet measured
experimentally. They are set to give the best argeement with experiments.
This would suggest that the $\beta$ order parameter would be positive,
in contrast to PC and PS headgroups, were negative signs were measured.
Even thought the signs turned out to be correct, the tested models would
not give a very good argeement with the experiments.
\begin{figure}[]
  \centering
  \includegraphics[width=9.0cm]{../Figs/HGorderparametersPG.pdf}
  \caption{\label{HGorderParametersPOPG}
    Order parameters for PG headgroup and glycerol backbone from simulations and experiments without CaCl$_2$ 
    (POPG from \cite{borle85} contains 10mM of PIPES, DPPG from \cite{wohlgemuth80} contains 10mM PIPES and 100mM CaCl,
    E.Coli PG results from \cite{gally81}).
    Signs are not known for experimental data. They are determined to give best agreement
    with simulations. This is not reliable and should be corrected when experimental
    data becomes available.
  }
\end{figure}

@@

\subsection{Lipid headgroup interactions in PC:PE and PC:PG mixtures}
The headgroup 
order parameters increase with the addition of negatively charged PS and PG lipids,
decrease when mixed with positively charged surfactants and
are less affected by the addition of zwitterionic PE lipids or cholesterol.
In addition to the results summarized in Fig. \ref{HGorderparametersPCvsPEPSPGchol},
also mixtures of PC with negatively charged PI, CL, PA, and zwitterionic SM
follow the eletrometer concept \cite{scherer87}.
to mixtures of differently charged
lipids are collected from different experiments in Fig. \ref{HGorderparametersPCvsPEPSPGchol}.
%\section{PC lipid headgroup response to different mixtures in experiments}
As shown in Fig. \ref{HGorderparametersPCvsPEPSPGchol}, order parameters of PC
headgroup behave in various lipid mixtures as expected from the electrometer concept \cite{seelig87, scherer87},
i.e., order parameters increase when anionic lipids are mixed with PC and decrease with cationic
surfactants. The changes with the addition of neutral lipids is significantly smaller.
%\begin{figure}[]
%  \centering
%  \includegraphics[width=8.0cm]{../Figs/HGorderparametersPCvsPEPSPGchol.eps}
%  \caption{\label{HGorderparametersPCvsPEPSPGchol}
%    PC headgroup order parameters from experiments of mixtures with
%    PE, PS, PG and cholesterol \cite{scherer87,scherer89,ferreira13}.
%    Signs are determined as discussed in \cite{botan15,ollila16}.
%  }
%\end{figure}

@@

\begin{figure*}[]
  \centering
  \includegraphics[width=16.0cm]{../Figs/HGorderparametersPCvsPEPG.eps}
  \caption{\label{HGorderparametersPCvsPEPSPG}
    PC headgroup order parameters from mixtures with PE, PS and PG
    lipids with various mole fractions from different simulation models and experiments \cite{scherer87}.
    Signs are determined as discussed in \cite{botan15,ollila16}.
  }
  \todo{Simulation of CHARMM36 at 298K should be maybe rerun with Gromacs 5.}
\end{figure*}

@@

\subsection{Ion binding to PE lipid bilayers}

\begin{figure}[]
  \centering
  \includegraphics[width=8.0cm]{../Figs/NAdensPE.eps}
  \caption{\label{NAdensPE}
    Sodium (solid line) and choride ion density profiles along membrane normal
    from different simulations with PE lipids.
  }
\end{figure}

@@

Comparison of Ca$2+$ binding in PG between CHARMM36 simulations and experiments \cite{borle85}
is shown in Fig. \ref{changesWITHCaClPG}. The decrease of $\alpha$ order parameter
is in agreement with experiments, while decerase of $\beta$ order parameter is overestimated.
The result is very similar to the results with PC in NMRlipids II publication \cite{catte16}.
It should be, however, noted that the $\beta$-order parameters are not actually measured for PG,
but they are calculated from empirical relation $\Delta S_{\beta}=0.43\Delta S_{\alpha}$ \cite{akutsu81}.
Anyway, the data presented in NMRlipids II project and in Fig. \ref{changesWITHCaClPG} together
suggest that Calcium binding is similarly overestimated by CHARMM36 model in pure POPC bilayers and mixtures with
POPG. The good agreement of $\alpha$ carbon would be explained by too weak dependence of its order
parameter of bound charge
\todo{The response of CHARMM36 to cationic surfactant against experiments \cite{scherer89} to be checked.
  I have already ran the simulations, analysis to be done.}.
\begin{figure*}[]
  \centering
  \includegraphics[width=18.0cm]{../Figs/CHANGESwithCaClPG.eps}
  \caption{\label{changesWITHCaClPG}
    PG order parameters as a function CaCl$_2$ concentration from experiments \cite{borle85} and CHARMM36 simulations.
    Note that beta order parameter is calculated from empirical relation $\Delta S_{\beta}=0.43\Delta S_{\alpha}$ \cite{akutsu81}, not actually measured. 
  }
\end{figure*}

@@

Also the experimental order parameters for PS and PG headgroups
as a function of CaCl$_2$ concentration are shown in Fig. \ref{PSPGchangesWITHCaCl}.
%\todo{These should be compared to simulations for potential structural interpretation of the changes.}
%\begin{figure}[]
%  \centering
%  \includegraphics[width=9.0cm]{../Figs/PGwithCaCl.eps}
%  \caption{\label{PSPGchangesWITHCaCl}
%    PG and PS order parameters a function CaCl$_2$ concentration taken from \cite{borle85} and \cite{roux90}, respectively.
%  }
%  \todo{Get the small concentration data from the inserts}
%\end{figure}
